Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-23 WSSSPE3/report/WSSSPE3_review.tex
Keynote: 3 models of sustainability

Turk then moved onto various different definitions of sustainability. His 
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software.
A funder Turk heard talked about sustainability as continuing to get funded. Turk also
recalled that Greg Wilson said in WSSSPE 1.1 that his view of sustainable software
was software that continued to give the same results over time. A last measure
of sustainability Turk presented was the ability to transition between different people
developing and using a piece of software.

>>>>

Turk then moved onto various different definitions of sustainability. His
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software. A
funder Turk heard talked about sustainability as continuing to get funded. Turk
also recalled that Greg Wilson. among others, said in WSSSPE1.1 that his view of
sustainable software was software that continued to give the same results over
time. A last measure of sustainability Turk presented was the ability to
transition between different people developing and using a piece of software.

At WSSSPE1, several different models were presented for ensuring sustainability.
Turk considered a familiar one was a funded piece of software where an external
agency provided funds to a group that was not necessarily exclusively working on
and developing a piece of software, keeps it going, and provides that out to the
scientific community. The model of productized software, in which a piece of
software has grown into something that individuals rely on to the point that
research groups or people are willing to support it at some pay rate, for
instance, subscription to use cloud services that deploy a piece of software, or
purchase of a piece of software. A last model Turk felt conflicted about is a
volunteer model, which is traditional old school---not modern-day open
source---developement.
